### 2022-11-17 chapter1.tex
Re-worded clustering algorithm discussion to apply to more generally and contrast previous methods to ours

\section*{Discussion}
% TODO: Highlight specific contributions
The orthologous groups and alignments yielded by this pipeline are a valuable resource for comparative studies of gene birth/death processes and protein evolution at the level of both entire proteomes and specific gene families in the \textit{Drosophila} genus. To assist these efforts, the final alignments after curation and the labels from the ``missing'' phylo-HMM are provided as supplemental data. Although this work focused on single copy orthologs, other studies may require different subsets of orthologous groups that demand other pre-processing and alignment strategies. Therefore, we have included the orthologous groups and the initial alignments with and without non-representative sequences in the supplemental data as well. While we anticipate these resources will remain relevant in the near term, the trends that permitted this work to substantially improve on previous efforts will render them obsolete in the coming years as more \textit{Drosophila} genomes are assembled and annotated. However, an authoritative and lasting set of orthologous groups and alignments is not the primary goal of this work. Instead, it serves as a case study in how dense taxonomic sampling and modern genome assembly and annotation pipelines present new opportunities and challenges to the traditional techniques for identifying and aligning orthologous groups.

For example, despite many additional pre-processing steps and other tweaks introduced by later authors, the basic framework of orthology inference by clustering the hit graph has remained largely unchanged in the past twenty years~\cite{Remm2001, Li2003, Jensen2007, Linard2011, Emms2015, Train2017, Cosentino2018}. This longevity is a testament to the robustness of the underlying idea that orthologous proteins should consistently identify each other as the most similar pairs between their genomes. Even as the number of genomes and their taxonomic density has increased dramatically, many orthology inference pipelines continue to use triangle clustering, which was originally applied to sets of far fewer and more distantly related genomes~\cite{Tatusov1997}. This mismatch in scale increases the chance of propagating annotation errors since only a small number of edges are needed to create or merge clusters. Fortunately, triangle clustering is a special case of the more general \textit{k}-clique percolation algorithm where \textit{k} is a tunable parameter that influences the tightness of a cluster. The optimal value of \textit{k} for a given set of genomes is unclear and likely depends on the desired trade-off between sensitivity and specificity. However, percolation theory or simulations may yield insights.

>>>>

\section*{Discussion}
The orthologous groups and alignments yielded by this pipeline are a valuable resource for comparative studies of gene birth/death processes and protein evolution at the level of both entire proteomes and specific gene families in the \textit{Drosophila} genus. To assist these efforts, the final alignments after curation and the labels from the ``missing'' phylo-HMM are provided as supplemental data. Although this work focused on single copy orthologs, other studies may require different subsets of orthologous groups that demand other pre-processing and alignment strategies. Therefore, we have included the orthologous groups and the initial alignments with and without non-representative sequences in the supplemental data as well. While we anticipate these resources will remain relevant in the near term, the trends that permitted this work to substantially improve on previous efforts will render them obsolete in the coming years as more \textit{Drosophila} genomes are assembled and annotated. However, an authoritative and lasting set of orthologous groups and alignments is not the primary goal of this work. Instead, it serves as a case study in how dense taxonomic sampling and modern genome assembly and annotation pipelines present new opportunities and challenges to the traditional techniques for identifying and aligning orthologous groups.

For example, despite many additional pre-processing steps and other tweaks introduced by later authors, the basic framework of orthology inference by clustering the hit graph has remained largely unchanged in the past twenty years~\cite{Tatusov1997, Remm2001, Li2003, Jensen2007, Linard2011, Emms2015, Train2017, Cosentino2018}. This longevity is a testament to the robustness of the underlying idea that orthologous proteins should consistently identify each other as the most similar pairs between their genomes. Even as the number of genomes and their taxonomic density has increased dramatically, many orthology inference pipelines continue to use algorithms which were originally applied to sets of far fewer and more distantly related genomes. This mismatch in scale increases the chance of propagating annotation errors since only a small number of edges are needed to create or merge clusters. Thus, we instead applied a generalization of the triangle and connected components clustering methods called \textit{k}-clique percolation where \textit{k} is a tunable parameter that influences the tightness of a cluster. The optimal value of \textit{k} for a given set of genomes is unclear and likely depends on the desired trade-off between sensitivity and specificity. However, percolation theory or simulations may yield insights.
